Modal Abstractions for Virtualizing Memory Addresses

### 2023-04-14 programlogic.tex
+

\begin{figure*}
  \begin{lstlisting}[language=Coq]
 Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\; \vpage$ : vProp $\Sigma$ := 
  $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(maddr l4e l3e l2e l1e) $\ast$ addr_L1$(\vaddr,\textsf{l1e}) \mapsto_{p}\{\mathsf{q}\} \vpage$.
 Definition L4_L1_PointsTo (maddr l4e l3e l2e l1e :word 64) : vProp $\Sigma$ :=
 $\lambda$ cr3val.
  $\ulcorner$ entry_present l4e $\land$ entry_present l3e $\land$ entry_present l2e $\land$ entry_present l1e$\urcorner$ $\ast$
  (l4M52 maddr cr3val) $\sim$ (l4off maddr cr3val) $\mapsto_{a}$ {q1}  l4e  $\ast$
  (l3M52 maddr l4e) $\sim$ (l3off maddr l4e)  $\mapsto_{a}$ {q2}  l3e $\ast$ 
  (l2M52 maddr l3e) $\sim$ (l2off maddr l3e) $\mapsto_{a}$ {q3}  l2e $\ast$
  (l1M52 maddr l2e) $\sim$ (l1off maddr l2e) $\mapsto_{a}$ {q4}  l1e.
\end{lstlisting}
\vspace{-1em}
\caption{A Strong Virtual Points-to Relation}
  \label{fig:strongvirtualpointsto}
\end{figure*}

@@

In Figure \ref{fig:addrswitch}, we see two different address-spaces rooted at R0 and R1 in the pre/post condition. To be put into words, while switching an address-space by executing \lstinline|mov_ctl|) updating the  \lstinline|cr3|, it is expected to mention resources of two different address-spaces in the specification. In the precondition,  \lstinline|cr3| is loaded with R0 and all the virtual-pointsto assertions $a_0 \mapsto_{\textsf{v}}p_0$ and $a_1 \mapsto_{\textsf{v}}p_1$ are in the current-view of the memory and valid in the ambient logic \textsf{vProp}. The other address-space with the root address R1 has its own resources ($b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$) which are invalid under the current address space rooted at R0. This pattern of specification requires resources being mentioned under the \textit{world} they are valid, i.e. a modal-framing within the specification. Once the current address-space becomes the one with root address R1, i.e.  \lstinline|cr3| is loaded with R1, the context-resources $b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$ are loaded into the current-view of the memory, and all the other resources that are valid under R0 are introduced into the modal context rooted at R0
\begin{figure}
\[
  \begin{array}{l}
    \hbox{(\TirNameStyle{ModalAddressSpace)}} \qquad
         [r](P) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{:=} \; \lambda \_,\; \mathsf{P}~\mathsf{r}. 
  \end{array}
  \]
  \label{fig:modaldef}
  \end{figure}

>>>>

\begin{figure*}
  \begin{lstlisting}[language=Coq]
 Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\; \vpage$ : vProp $\Sigma$ := 
  $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(maddr l4e l3e l2e l1e) $\ast$ addr_L1$(\vaddr,\textsf{l1e}) \mapsto_{p}\{\mathsf{q}\} \vpage$.
 Definition L4_L1_PointsTo (maddr l4e l3e l2e l1e :word 64) : vProp $\Sigma$ :=
 $\lambda$ cr3val.
  $\ulcorner$ entry_present l4e $\land$ entry_present l3e $\land$ entry_present l2e $\land$ entry_present l1e$\urcorner$ $\ast$
  (l4M52 maddr cr3val) $\sim$ (l4off maddr cr3val) $\mapsto_{a}$ {q1}  l4e  $\ast$
  (l3M52 maddr l4e) $\sim$ (l3off maddr l4e)  $\mapsto_{a}$ {q2}  l3e $\ast$ 
  (l2M52 maddr l3e) $\sim$ (l2off maddr l3e) $\mapsto_{a}$ {q3}  l2e $\ast$
  (l1M52 maddr l2e) $\sim$ (l1off maddr l2e) $\mapsto_{a}$ {q4}  l1e.
\end{lstlisting}
\vspace{-1em}
\caption{A Strong Virtual Points-to Relation}
\todo[inline,color=red]{Colin, I updated the definition with entry present}
  \label{fig:strongvirtualpointsto}
\end{figure*}

@@

In Figure \ref{fig:addrswitch}, we see two different address-spaces rooted at R0 and R1 in the pre/post condition. To be put into words, while switching an address-space by executing \lstinline|mov_ctl|) updating the  \lstinline|cr3|, it is expected to mention resources of two different address-spaces in the specification. In the precondition,  \lstinline|cr3| is loaded with R0 and all the virtual-pointsto assertions $a_0 \mapsto_{\textsf{v}}p_0$ and $a_1 \mapsto_{\textsf{v}}p_1$ are in the current-view of the memory and valid in the ambient logic \textsf{vProp}. The other address-space with the root address R1 has its own resources ($b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$) which are invalid under the current address space rooted at R0. This pattern of specification requires resources being mentioned under the \textit{world} they are valid, i.e. a modal-framing within the specification. Once the current address-space becomes the one with root address R1, i.e.  \lstinline|cr3| is loaded with R1, the context-resources $b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$ are loaded into the current-view of the memory, and all the other resources that are valid under R0 are introduced into the modal context rooted at R0
\begin{figure}
\[
  \begin{array}{l}
    \hbox{(\TirNameStyle{ModalAddressSpace})} \qquad
         [r](\mathsf{P}) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{=} \; \lambda \_,\; \mathsf{P}~\mathsf{r}.\\
    \hbox{(\TirNameStyle{ModalAddressSpaceMono})} \qquad (\mathsf{P} \vdash \mathsf{Q}) \vdash  ([r](\mathsf{P}) \vdash  [r](\mathsf{Q}))\\
    \hbox{(\TirNameStyle{ModalAddressSpaceSep})} \qquad [r](\mathsf{P} \ast \mathsf{Q}) \dashv\vdash ([r](\mathsf{P}) \ast [r](\mathsf{Q})\\
    \hbox{(\TirNameStyle{ModalAddressSpacePure})} \qquad  \mathsf{pure\_vprop P}(\mathsf{iPropI }\Sigma)\textsf{ : vProp }\Sigma \stackrel{\triangle}{=} \; \lambda \_ \ldotp \mathsf{P}\\
    \hbox{(\TirNameStyle{ModalAddressSpaceFact})} \qquad  \mathsf{Fact P}(\mathsf{vProp }\Sigma) \stackrel{\triangle}{=} \;\forall \; ,r \; r' \ldotp  \mathsf{P r} \dashv\vdash \mathsf{P r'}\\
    \hbox{(\TirNameStyle{ModalAddressSpaceFactIntroElim})} \\
    \qquad\qquad \mathsf{Fact pure\_vprop P} \vdash  ([r](\mathsf{pure\_vprop P}) \dashv\vdash (\mathsf{pure\_vprop P}) (\mathsf{vProp }\Sigma))\\

    \hbox{(\TirNameStyle{ModalAddressSpaceAnd})} \qquad   [r](\mathsf{P}\land \mathsf{Q}) \dashv\vdash [r](\mathsf{P})\land [r](\mathsf{Q})\\    
    \hbox{(\TirNameStyle{ModalAddressSpaceOr})} \qquad    [r](\mathsf{P}\lor \mathsf{Q}) \dashv\vdash [r](\mathsf{P})\lor [r](\mathsf{Q})\\   
    \hbox{(\TirNameStyle{ModalAddressSpaceFactIntroElim})} \\
    \qquad\qquad \mathsf{Fact pure\_vprop P} \rightarrow  ([r](\mathsf{pure\_vprop P}) \dashv\vdash (\mathsf{pure\_vprop P}) (\mathsf{vProp }\Sigma))\\

  \end{array}
  \]
  \label{fig:modaldef}
  \end{figure}
